sentiment-analysis

### 2016-08-03 eda/sentiment-analysis.Rmd
cleaning data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(jsonlite)
library(stringr)
library(ggplot2)
library(dplyr)
library(tidytext)
library(tm)
```

@@

Now to apply this sentiment analysis to the Yelp review dataset, I have to mutate the data so that each word for every review exists in one row. I also need to remove stop words like "the", "and", "I", etc. since they don't indicate anything about the attitude of the reviewer and shouldn't affect the valence score. A library called __tm__ for text mining can eliminate stop words. 
```{r, eval = F}
review_words <- reviews %>% 
  mutate(words = str_split(text, "[ ,.]")) %>% 
  unnest(words)

@@

```{r, eval = F}
# unnest_tokens splits column into tokens using tokenizers package
# splits "text" column into individual words to analyze sentiment for further
# exploration
reviews_text <- reviews %>% 
  select(business_id, review_id, stars, text) %>%
  unnest_tokens(word, text) %>% 
  anti_join(stop_words) %>% 
  arrange(review_id)

# rank words from -5 (negativity) to 5 (positivity) using AFINN lexicon
# sentiments: 
#   data frame with 4 variables: word, sentiment, lexicon, score
# lexicon can be "nrc", "bing", or "AFINN"
valence_score <- sentiments %>% 
  filter(lexicon == "AFINN") %>% 
  select(word, score)

# calculate mean for each group of stars
review_scores <- reviews_text %>% 
  inner_join(valence_score, by = "word") %>% 
  group_by(review_id, stars) %>% 
  summarise(sentiment = mean(score))
```

@@

```{r, eval = F}
# boxplot for each group
ggplot(review_scores) +
  geom_boxplot(aes(stars, sentiment, group = stars)) +
  labs(x = "Business Star Rating",
       y = "Mean Sentiment Score",
       title = "Mean Sentiment Score vs. Business Star Rating") +
  theme_bw()
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(jsonlite)
library(stringr)
library(ggplot2)
library(dplyr)
library(tm)
```

@@

I'll read in the first 500,000 observations of the Yelp reviewers data, which has been written to a csv file.
```{r}
reviews <- read_csv("../clean_data/reviews.csv")
```

Now to apply this sentiment analysis to the Yelp review dataset, I have to mutate the data so that each word for every review exists in one row. I also need to remove stop words like "the", "and", "I", etc. since they don't indicate anything about the attitude of the reviewer and shouldn't affect the valence score. A library called __tm__ for text mining can eliminate stop words.

```{r, eval = F}
review_words <- reviews %>% 
  mutate(words = str_split(text, "[ ,.]")) %>% 
  unnest(words)

@@

```{r, eval = T}
# moved all tidy procedures to CleanData.R
review_scores <- read_csv("../clean_data/review_scores.csv")
```

@@

```{r, eval = T}
# boxplot for each group
ggplot(review_scores) +
  geom_boxplot(aes(stars, sentiment, group = stars)) +
  labs(x = "Business Star Rating",
       y = "Mean Sentiment Score",
       title = "Mean Sentiment Score vs. Business Star Rating") +
  theme_bw()
```

Positive correlation exists! Some inaccuracy (outliers on boxplots). Find out most common positive and negative words in reviews, creating a summary of the counts.

```{r fig.width = 8}
summaries <- read_csv("../clean_data/summaries.csv")

summaries %>% 
  ggplot(aes(reviews, average_star, color = score)) +
  geom_point() +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1, hjust = 1) +
  scale_x_log10() +
  geom_hline(yintercept = mean(reviews$stars), color = "black", lty = 2) +
  scale_color_gradient2(low = "red", high = "blue", midpoint = 0, mid = "gray") +
  labs(x = "# of reviews",
       y = "Average Stars",
       color = "AFINN") +
  theme_bw()
```
